---
title: "homework-week2"
author: "lewris Mota"
date: "February 10, 2019"
output:
  html_document: default
  pdf_document: default
---

####**Choose six recent popular movies.  Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5.  Take the results (observations) and store them in a SQL database.  Load the information into an R dataframe.**




```{r , include=FALSE}

library(RMySQL)

```



```{r}
conexion <- dbConnect(MySQL(),user="root",password="root", 
                      dbname="movies", host = "127.0.0.1", port=3306)
```

####_Tables of Movies Database_
```{r}

dbListTables(conexion)

```

####_ Load data into the dataframe
```{r , warning=FALSE}

result_set <- dbSendQuery(conexion,"select movie_name, movie_rating from movies m inner join ratings s on s.movie_id = m.movie_id ")

movies_dt <- fetch(result_set)

movies_dt

```

#### Now lets get the average rating for movies

```{r , warning=FALSE}

result_set <- dbSendQuery(conexion,"select movie_name, AVG(movie_rating) as 'average_rating' from movies m inner join ratings s on s.movie_id = m.movie_id group by movie_name")

movies_dt <- fetch(result_set)

movies_dt

```


