---
title: "week-3-assig"
author: "Lewris Mota"
date: "February 15, 2019"
output: html_document
---
```{r echo=FALSE}

library(stringr)

```


####**Copy the introductory example. The vector name stores the extracted names.**
```{r}
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555
-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson,Homer5553642Dr. Julius Hibbert"
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
```


**(a) Use the tools of this chapter to rearrange the vector so that all elements conform to
the standard first_name last_name.**
```{r}

standard <- paste(str_extract(name,"[^DrRev][:alpha:]{1,}"),str_extract(name,"(?<=[:blank:]|,)[a-zA-Z]{1,}$"))
standard
```


**(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr. ).**
```{r}

has_tittle <- str_detect(name,"(^[a-zA-Z]{1,}\\.)")
has_tittle
#titles
unlist(str_extract_all(name,"(^[a-zA-Z]{1,}\\.)"))
#vector type
typeof(has_tittle)


```



**(c) Construct a logical vector indicating whether a character has a second name.**
```{r}

has_2name <- str_detect(name,"[a-zA-Z]{1,}\\.\\s(?=([a-zA-Z]{1,})$)")

#extract second name
unlist(str_extract_all(name,"[a-zA-Z]{1,}\\.\\s(?=([a-zA-Z]{1,})$)"))

typeof(has_2name)

```


####**Describe the types of strings that conform to the following regular expressions and
construct an example that is matched by the regular expression.**

**(a) [0-9]+\\$**

Matches a string that has one or more numeric characters followed by a $ sign in that same string
```{r}
#true if if numbers and $ are present in the same string one after the other
str_detect("6685$6","[0-9]+\\$")

#false if the line does not have numbers
str_detect("$","[0-9]+\\$")

#false if the line does not have $ sign
str_detect("6685","[0-9]+\\$")

```

**(b)\\b[a-z]{1,4}\\b **


Matches any lower case strings that have from 1 to 4 letters/characters

```{r}
#returns all the matches that have less than 4 lowercase characters
str_extract_all("I Am trying this reg Expression","\\b[a-z]{1,4}\\b")

```


(c).*?\\.txt$

This expression reads a complete string and returns a match if the last string at the end of the line has the .txt characters. 
```{r}
#true if at the end
str_detect("I am lewris.txt",".*?\\.txt$")
#false if the line does not end with .txt
str_detect("I am lewris.txt but not at the end",".*?\\.txt$")


```

**(d) \\d{2}/\\d{2}/\\d{4}**

Matches any occurence of dates in formats mm/dd/yyyy or dd/mm/yyyy tha appear in any string.
```{r}

str_extract_all("5505/06/1886, 10/06/1885, 1/2/2016  ","\\d{2}/\\d{2}/\\d{4}")



```


**(e) <(.+?)>.+?</\\1>**

Matches any occurence of a strings properly formated as html tags with opening and closing tags that are not empty. The content between html tags is returned even if more html tags are present.
```{r}

str_extract_all("<empty></empty>><tittle>aaaa</tittle>","<(.+?)>.+?</\\1>")
#tags inside tags are also returned
str_extract_all("<empty></empty>><tittle>aaaa</tittle><body><h1>trying html tags</h1></body>","<(.+?)>.+?</\\1>")

```


