---
title: "Week8"
author: "Lewris Mota"
date: "March 31, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jsonlite)
library(tidyverse)
library(RCurl)
library(kableExtra)
```

#### Instructions

The New York Times web site provides a rich set of APIs, as described here: http://developer.nytimes.com/docs
You’ll need to start by signing up for an API key.
Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and
transform it to an R dataframe.

#### Load API data

Data is loaded from the web api. The information being read are the 20 best sellers books registered by NYtimes api.
```{r}

link2 <- "https://api.nytimes.com/svc/books/v3/lists/best-sellers/history.json?api-key=G1cUkg2rCJr5xWog5sRRnePn1JTtUBCZ"

booksDT <- getURL(link2)

```

#### Reading Data

The data is read and converted from json to a list.
```{r}

bookl <- booksDT %>% fromJSON(simplifyVector = FALSE)
bookList <- bookl$results 
```

#### Preparing Dataframe
```{r}

#getting the headers names from list name
headers <- bookList[[1]] %>% unlist() %>% names()
headers <- headers[1:8] %>% as.list()

#empty dataframe for storing json data. This dataframe will have 8 columns.
books <- data.frame(matrix(ncol = 8))
names(books) <- headers


#data is processed and stored in the dataframe
for (i in 1:20) {
  bo <- bookList[[i]][1:8] %>% unlist() %>% as.vector()
  books <- rbind(books,bo)
}

```

#### Data Display


```{r}
books %>% na.omit %>% kable() %>% 
kable_styling()
```

